#Practical Machine learning by Johns hopkins Prediction Assignment WriteUp Report

##Link Git Hub repo

###How Model has been constructed

####First lets read the data

```{r}

library(modeest)
library(caret)
require(data.table)

trainingInputData<-read.csv("F:/jeevi/AICourses/writeUP/pml-training.csv", na.strings=c("NA",""))
testingInputData<-read.csv("F:/jeevi/AICourses/writeUP/pml-testing.csv", na.strings=c("NA",""))
inTrain <- createDataPartition(trainingInputData$classe, p=0.75, list = FALSE)
trainSet <- trainingInputData[inTrain,]
pretestSet <- trainingInputData[-inTrain,]
```

####Data Cleaning Techniques
Here I'm applying two step process  
1. Columns like names, rowID, datatime stamps are not adding any values to the prediction process. Hence forth removing the same.  
2. Remove NA columns which are having more than 10000 NA values.    
     Reason : Because if we predict using KNN impute method when there are more than half of the values are NA. Prediction process won't be                         reliable.
     
```{r}
trainSetRemovedColumns<-trainSet[, -(1:6)]
pretestSetRemovedColumns<-pretestSet[, -(1:6)]
testSetRemovedColumns<-testingInputData[, -(1:6)]

colNA<-which(colSums(is.na(trainSetRemovedColumns)) > 10000)

trainSetCleaned<-trainSetRemovedColumns[, -colNA]
pretestSetRemovedColumns<-pretestSetRemovedColumns[, -colNA]
testSetCleaned<-testSetRemovedColumns[, -colNA]

# Remove Class variables
testSetWithoutClassVariable<-trainSetCleaned[, -54]
pretestSetWithoutClassVariable<-pretestSetRemovedColumns[, -54]
```
Applying GBM method to train the system  
Note: Other methods were applied like Random forest and LDA.   
All those method returned nearly same accuracy. You can find those method step process at the end of the document.   
In appendix 1 and appendix 2 respectively.  

```{r}
ModelGBM <- train(trainSetCleaned$classe~ .,method="gbm", data=testSetWithoutClassVariable)
```
Note: As there was limit to the charaters(i.e., 2000) in the report i have manually edited HTML file and removed the content after iteration 2.  
Lets analyse trained set  
In sample error: Error which comes out from training.  
OutSample error: Error Which is from new Data Set.  
Here Validation set is used to measure Out Sample error on Trained set  
```{R}
predictTrain<-predict(ModelGBM, testSetWithoutClassVariable)
predictPreTest<-predict(ModelGBM, pretestSetWithoutClassVariable)
predictTest<-predict(ModelGBM, testSetCleaned)


tableTrain<-table(predictTrain, trainSetCleaned$classe)
tablePretest<-table(predictPreTest, pretestSetRemovedColumns$classe)

confusionMatrix(tableTrain)
```
IN sample Accuracy is 99.25 IN Sample error is 0.75%  
**Do the authors describe what they expect the out of sample error to be and estimate the error appropriately with cross-validation?**    
*Now check pretext or validation test to find OUT sample error. OUT Sample error is calculated on Data which was not used in training the model.*  
**
Out Sample Accuracy is :98.9%  
Out Sample error Rate is : 1.1% **  
Model Prediction on test Set  
```{r}
predictTest
```
**END OF REPORT**  

#Appendix 1:    
**Train the model using random forest**  
Creating data set remain same as we followed in the process GBM  
```{r}
ModelRF <- train(trainSetCleaned$classe~ .,method="rf", data=testSetWithoutClassVariable)

```

```{r}
predictTrain<-predict(ModelRF, testSetWithoutClassVariable)
predictPreTest<-predict(ModelRF, pretestSetWithoutClassVariable)
predictTest<-predict(ModelRF, testSetCleaned)

 
tableTrain<-table(predictTrain, trainSetCleaned$classe)
tablePretest<-table(predictPreTest, pretestSetRemovedColumns$classe)
```
IN sample Accuracy is 100% IN Sample error is 0%  
```{r}
confusionMatrix(tableTrain)
```
**Out sample Accuracy is 99.79 IN Sample error is 0.21%**  

```{r}
confusionMatrix(tablePretest)[3]
```
Prediction on Test Set  
```{r}
predictTest
```

#Appendix 2:  
**Train the model USING LDA**  
```{r}
Modellda <- train(trainSetCleaned$classe~ .,method="lda", data=testSetWithoutClassVariable)
predictTrain<-predict(ModelGBM, testSetWithoutClassVariable)
predictPreTest<-predict(ModelGBM, pretestSetWithoutClassVariable)
predictTest<-predict(ModelGBM, testSetCleaned)

tableTrain<-table(predictTrain, trainSetCleaned$classe)
tablePretest<-table(predictPreTest, pretestSetRemovedColumns$classe)
confusionMatrix(tableTrain)
```
IN sample Accuracy is 100% IN Sample error is 0%  
Out sample Accuracy is 99.85 IN Sample error is 0.15%  

```{r}
confusionMatrix(tablePretest)[3]
```
**Prediction on Test Set**  
```{r}
predictTest
```
